---
title: "Untitled"
author: "Segovia D L Revilla, Lucia"
date: "07/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Selecting food items to be included in the review

## Getting the data

1. Download food supply data: The data can be downloaded from [FAOSTAT website](http://www.fao.org/faostat/en/). On "Data" tab, we accessed "Food Balances (2010-)" (under Food Balances). We selected:

  - "Regions (FAO)" = Eastern Africa + (Total), Middle Africa + (Total),  Southern Africa + (Total), Western Africa + (Total).
  - "Elements" = Food Supply quantity (kg/capita/yr), Food supply (kcal/capita/year)
  - "Items" = "Select All" but "Population".
  - "Years" = "Select All". (Which included from 2010-2018)

Last accessed and downloading date: 25/04/2022.


2. Download food composition data: The data can be downloaded from Joy et al. (2014) (https://onlinelibrary.wiley.com/doi/full/10.1111/ppl.12144). Supporting Information: Table S2. Food mineral composition data from literature sources, used in conjunction with FBSs to estimate dietary mineral availability. File name: ppl12144-sup-0002-TableS2.xlsx. 

Last accessed and downloading on: 25/04/2022.

In previous work by Joy et al. (2014), they used Western FCT for the Middle region as well, and that is what we have done. *This is to be reviewed* 



```{r cars}
summary(cars)
```

## Matching food supply and food composition data

Regional FCT were prepared to be matched with "old Food Balance Sheet", however we are using the "new" Food Balance Sheet. There methodological differences that need to be explored if we want to combine and use data from both (new and old), but that is beyond the scope of this review. Nevertheless, there is one arrange to be done. Rice and Groundnuts code have been change. This is because the reported commodity state is different (husked vs paddy and shelled vs in shell). We are not making any adjustments to this as we do are not using the data for analysis but for rough estimations of the importance of food items in supply. 

# Limitations

- Food Composition data:

1. The selection of the food composition data used to determine the energy and mineral supply have an impact on the food list that will be selected. For instance, if we use the Southern Africa region FCT to be the Middle region FCT, instead of the Western region, there are two items that will not be selected "sesamee seeds" and "sweet potatoes", while "plantain" will be included. 

Items that were not reported in the FCT, and as we can see in the table below, only "miscellaneous" was higher than zero.

`r not_reported %>% mutate_at("mean", ~round(., 2)) %>% arrange(desc(mean)) %>% gt()`. 


#Checking with energy supply

For aggregated data at region level from 2014-2018. The mean supply of energy was calculated and selected the top ten food categories. Then compared with our results. Four food categories were reported as inconsistent bewteen the FAO and our calculated top ten energy contributors. "milk - excluding butter", "millet and products" were reported as top10 suppliers of Energy for the region whereas we found groundnuts and plantains. However, those two food categories were reported in the final food list, leading to no discrepancies.        

```{r }

setdiff(tolower(ener_list2),ener_list1)
setdiff(ener_list1, tolower(ener_list2))

```

#Identifying FBS categories that needed further information

#1) Food Balance Sheets

The FBS categories that needed further information were:

13) 2605	Vegetables, Other
15) 2534	Roots, Other
19) 2549	Pulses, Other and products
20  2570	Oilcrops Oil, Other
22) 2657	Beverages, Fermented
24) 2736	Offals, Edible

Relation between FBS and SUA food categories can be found in Related Documents
FBS and SUA list [http://fenixservices.fao.org/faostat/static/documents/SCL/FBS%20and%20SUA%20list.xlsx]

#2) Supply and Utilization Accounts

http://www.fao.org/faostat/en/#data/SCL

Regions (ISO3) -> Select:

-- Eastern Africa > (List)
-- Middle Africa > (List)
-- Southern Africa > (List)
-- Western Africa > (List)

Elements -> Select:

Food supply (kcal/capita/day)
Food suply quantity (g/capita/day)

Items (CPC) -> Select:

Crops Primary > (List) -> Oilcrops, Oil Equivalent > (List)
Crops Primary > (List) -> Pulses > (List)
Crops Primary > (List) -> Roots and Tubers > (List)
Crops Primary > (List) -> Vegetables > (List)
Livestock primary > (List) -> Offals Edible Primary > (List)

Years -> Select All

Output Type = Table, File Type = csv, Output Formatting Options = c(Flags, Codes, Units)

Last Update: March 19, 2021

Downloaded on 28/07/2021 and stored as SUA-others_SSA_2014-2018.csv

##Beverages, fermented

http://www.fao.org/faostat/en/#data/SCL

Regions (ISO3) -> Select:

-- Eastern Africa > (List)
-- Middle Africa > (List)
-- Southern Africa > (List)
-- Western Africa > (List)

Elements -> Select:

Food supply (kcal/capita/day)
Food suply quantity (g/capita/day)

Items (CPC) -> Select:

Grand Total > (List) -> Beer of barley
Grand Total > (List) -> Beer of maize
Grand Total > (List) -> Beer of millet
Grand Total > (List) -> Beer of sorghum
Grand Total > (List) -> Beverage, fermented rice
Grand Total > (List) -> Beverage, fermented wheat

Years -> Select All

Output Type = Table, File Type = csv, Output Formatting Options = c(Flags, Codes, Units)

Last Update: March 19, 2021

Downloaded on 05/08/2021 and stored as SUA-others_SSA_2014-2018.csv


We have removed grouped items, because they provided no better information that what we have, and 
we have removed those items that were already covered in the FBS list and/or were left out in the previous selection step. I.e., for vegetables, onions and tomatoes are reported as independent group of items in FBS, however they were not included following the previously established criteria (See above).

Hence, we filter out (using the fooditem.code)

- "01253.02" - tomatoes, 
- "01234" - onions,
- "01290.90" - vegetables, fresh, nes,
- "01520.01" - cassava

i.	2534 - Roots, Other: 01550 - Taro (cocoyam)   
ii.	2549 - Pulses, Other and products: 01706 - Cow peas, dry
iii.	2605 - Vegetables, Other: 01212 - Cabbages and other brassicas
iv.	2570 - Oilcrops Oil, Other: 01449.01 - Melonseed
v.	2657 - Beverages, Fermented: None 
vi.	2736 - Offals, Edible: 21151 - Offals, edible, cattle


Top 3 vegetables, other by quantity are "cabbages and other brassicas", carrots and turnips and chilies!

Energy-wise:

 01290.01        Maize, green                   340
 2 01212         Cabbages and other brassicas   250
 3 01235         Pumpkins, squash and gourds    222

Qty-wise:

   01212         Cabbages and other brassicas   613
 2 01251         Carrots and turnips            405
 3 01231         Chillies and peppers, green    360
 

ISSUE with Dictionary --> SUA and Fish share the same code!!

 ID_0  FoodName_0        ID_1 FoodName_1                  ID_2  FoodName_2                  
  <chr> <chr>            <dbl> <chr>                       <chr> <chr>                       
1 RT    Roots and Tubers  2531 potatoes and products       1510  potatoes                    
2 RT    Roots and Tubers  2533 sweet potatoes and products 1530  sweet potatoes              
3 AP    Animal Products   2782 fish, liver oil             1510  freshwater fish, liver oil  
4 AP    Animal Products   2763 pelagic fish                1530  pelagic fish, frozen, fillet

#Fish

FAO. 2021. Fishery and Aquaculture Statistics. Global production by production source 1950-2019 (FishstatJ). In: FAO Fisheries Division [online]. Rome. Updated 2021. www.fao.org/fishery/statistics/software/fishstatj/en

Global production by production source 2014-2018. We added the FAOSTAT groups to be able to group by that categories and identify the fishes of the categories that we are interested in: Freshwater, Pelagic and Demersal. We have summarized by continent having previously selected the data only from the four region of interest (Western, Middle, Eastern and Southern Africa).


Preferences (manage visible column (F6)) -> Dimension -> ASFIS species -> New... 
FAOSTAT group of species -> Apply and Close. 

